Noncoherent least squares estimators of carrier phase and amplitude

### 2012-10-23 other/confnoncoherent/paper.tex
added stuff for silly new icassp fromat

%NOTE: If you use the natbib package, you should use natbibspacing.sty instead, and be sure to load it after natbib:
\usepackage[square,comma,numbers,sort&compress]{natbib}
\usepackage{natbibspacing}
\setlength{\bibspacing}{0cm}

@@

%Before describing the statistical properties of the least squares estimator, we first require some properties of complex valued random variables.  
A complex random variable $X$ is said to be \emph{circularly symmetric} if the distribution of its phase $\angle{X}$ is uniform on $[0,2\pi)$ and is independent of the distribution of its magnitude $\abs{X}$.  That is, if $Z \geq 0$ and $\Theta \in [0,2\pi)$ are real random variables such that $Ze^{j\Theta} = X$, then $\Theta$ is uniformly distributed on $[0,2\pi)$ and is independent of $Z$.  If the probability density function (pdf) of $Z$ is $f_Z(z)$, then the joint pdf of $\Theta$ and $Z$ is 
\[
f_{Z,\Theta}(z,\theta) = \frac{1}{2\pi}f_Z(z).
\]  
If $X$ is circularly symmetric, then for any real number $\phi$, the distribution of $X$ is the same as that of $e^{j\phi}X$.  %If $\expect\abs{X} = \expect Z$ is finite, then $X$ has zero mean because
% \begin{align*}
% \expect X &= \int_{0}^{2\pi} \int_{0}^\infty z e^{j\theta} f_{Z,\Theta}(z,\theta) dz d\theta \\
% &= \frac{1}{2\pi} \int_{0}^{2\pi} e^{j\theta} \int_{0}^\infty z f_Z(z) dz d\theta \\
% &= \frac{1}{2\pi}\expect Z \int_{0}^{2\pi} e^{j\theta} d\theta = 0.
% \end{align*}

We will have particular use of complex random variables of the form $1 + X$ where $X$ is circularly symmetric.  Let $R \geq 0$ and $\Phi \in [0,2\pi)$ be real random variables satisfying, 
\[
R e^{j\Phi} = 1 + X.
\]
The joint distribution of $R$ and $\Phi$ can be shown to be
\[
f(r,\phi) = \frac{r f_Z(\sqrt{r^2 - 2r\cos\phi + 1})}{2\pi\sqrt{r^2 - 2r\cos\phi + 1}}.
\]

@@

We have studied the least squares estimator of carrier phase and amplitude from the observation of $L$ noisy $M$-PSK symbols.  The estimator can be computed in $O(L\log L)$ operations using the algorithm of Mackenthun~\cite{Mackenthun1994}, and is the maximum likelihood estimator in the case that the noise is additive white and Gaussian.  We show that the phase estimator $\hat{\theta}$ is strongly consistent and asymptotically normally distributed.  However, the amplitude estimator $\hat{\rho}$ is biased, and converges to $G(0)\rho_0$.  This bias is large when the signal to noise ratio is small.  It would be interesting to investigate methods for correcting this bias.

\begin{figure}[p]
	\centering
		\includegraphics[width=\linewidth]{code/data/plotncM2-2.mps}
		\caption{Phase error for BPSK}
		\label{fig:plotphaseM2}
\end{figure}

@@

\begin{figure}[p]
	\centering
		\includegraphics[width=\linewidth]{code/data/plotncM2-1.mps}
		\caption{Amplitude error for BPSK}
		\label{fig:plotampM2}
\end{figure}

@@

\small
\bibliography{bib}

>>>>

%NOTE: If you use the natbib package, you should use natbibspacing.sty instead, and be sure to load it after natbib:
\usepackage[square,comma,numbers,sort&compress]{natbib}
\usepackage{bibspacing}
\setlength{\bibspacing}{0cm}

@@

Before describing the statistical properties of the least squares estimator, we first require some properties of complex valued random variables.  
A complex random variable $X$ is said to be \emph{circularly symmetric} if the distribution of its phase $\angle{X}$ is uniform on $[0,2\pi)$ and is independent of the distribution of its magnitude $\abs{X}$.  That is, if $Z \geq 0$ and $\Theta \in [0,2\pi)$ are real random variables such that $Ze^{j\Theta} = X$, then $\Theta$ is uniformly distributed on $[0,2\pi)$ and is independent of $Z$.  If the probability density function (pdf) of $Z$ is $f_Z(z)$, then the joint pdf of $\Theta$ and $Z$ is 
\[
f_{Z,\Theta}(z,\theta) = \frac{1}{2\pi}f_Z(z).
\]  
If $X$ is circularly symmetric, then for any real number $\phi$, the distribution of $X$ is the same as that of $e^{j\phi}X$.  %If $\expect\abs{X} = \expect Z$ is finite, then $X$ has zero mean because
% \begin{align*}
% \expect X &= \int_{0}^{2\pi} \int_{0}^\infty z e^{j\theta} f_{Z,\Theta}(z,\theta) dz d\theta \\
% &= \frac{1}{2\pi} \int_{0}^{2\pi} e^{j\theta} \int_{0}^\infty z f_Z(z) dz d\theta \\
% &= \frac{1}{2\pi}\expect Z \int_{0}^{2\pi} e^{j\theta} d\theta = 0.
% \end{align*}

% We will have particular use of complex random variables of the form $1 + X$ where $X$ is circularly symmetric.  Let $R \geq 0$ and $\Phi \in [0,2\pi)$ be real random variables satisfying, 
% \[
% R e^{j\Phi} = 1 + X.
% \]
% The joint distribution of $R$ and $\Phi$ can be shown to be
% \[
% f(r,\phi) = \frac{r f_Z(\sqrt{r^2 - 2r\cos\phi + 1})}{2\pi\sqrt{r^2 - 2r\cos\phi + 1}}.
% \]

@@

The problem of estimating carrier phase and amplitude from the observation of $L$ noisy $M$-PSK symbols has undergone extensive prior study~\cite{ViterbiViterbi_phase_est_1983,Cowley_ref_sym_carr_1998,Wilson1989,Makrakis1990,Liu1991,Mackenthun1994,Sweldens2001,McKilliamLinearTimeBlockPSK2009,Divsalar1990}.  The least squares estimator can be computed in $O(L\log L)$ operations using the algorithm of Mackenthun~\cite{Mackenthun1994}, and this is the maximum likelihood estimator in the case that the noise is additive white and Gaussian.  The algorithm was later rediscovered by Sweldens~\cite{Sweldens2001}.  In this paper we described the statistical properties of the least squares estimator. We showed that the phase estimator $\hat{\theta}$ is strongly consistent and asymptotically normally distributed.  However, the amplitude estimator $\hat{\rho}$ is biased, and converges to $G(0)\rho_0$.  This bias is large when the signal to noise ratio is small.  It would be interesting to investigate methods for correcting this bias.

%\newpage
\begin{figure}[p]
	\centering
		\includegraphics[width=\linewidth]{code/data/plotncM2-2.mps}
		\caption{Phase error for BPSK}
		\label{fig:plotphaseM2}
\end{figure}

@@

%\pagebreak
%\newpage
\begin{figure}[p]
	\centering
		\includegraphics[width=\linewidth]{code/data/plotncM2-1.mps}
		\caption{Amplitude error for BPSK}
		\label{fig:plotampM2}
\end{figure}

@@

%\vfill\pagebreak
\vfill\pagebreak
%\newpage
%\small
\bibliography{bib}
